### 2016-10-14 ms_disp1/ms_disp1.tex
updated text

\documentclass[aps,prd,twocolumn,superscriptaddress,amsfont,amssymb,amsmath,nofootinbib,showpacs,balancelastpage]{revtex4-1}
%\documentclass[prd,superscriptaddress,twocolumn,floatfix]{revtex4}

\usepackage{graphicx,longtable,natbib,mathrsfs,color}
\usepackage{txfonts,aas_macros}

\newcommand{\bs}{\boldsymbol}
\newcommand{\diff}{{\mathrm d}}
\newcommand{\Cov}{\mathsf{C}}
\newcommand{\Fish}{\mathsf{F}}
\newcommand{\Cl}{\mathcal {C}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\Msun}{M_\odot}
\newcommand{\lb}{\left\langle}
\newcommand{\rb}{\right\rangle}
\newcommand{\tcb}{\textcolor{blue}}
\newcommand{\tcr}{\textcolor{red}}

\begin{document}

\addtolength{\hoffset}{-0.525cm}
\addtolength{\textwidth}{1.05cm}
\title{Lagrangian Space Nonlinear $E$-mode clustering}

\author{Hao-Ran~Yu}%\email{haoran@cita.utoronto.ca}
\affiliation{Kavli Institute for Astronomy and Astrophysics, Peking University, Beijing 100871, China}
\affiliation{Canadian Institute for Theoretical Astrophysics, University of Toronto, 60 St. George Street, Toronto, Ontario M5S 3H8, Canada}

\author{Ue-Li~Pen}%\email{pen@cita.utoronto.ca}
\affiliation{Canadian Institute for Theoretical Astrophysics, University of Toronto, 60 St. George Street, Toronto, Ontario M5S 3H8, Canada}
\affiliation{Dunlap Institute for Astronomy and Astrophysics, University of Toronto, 50 St. George Street, Toronto, Ontario M5S 3H4, Canada}
\affiliation{Canadian Institute for Advanced Research, CIFAR Program in Gravitation and Cosmology, Toronto, Ontario M5G 1Z8, Canada}
\affiliation{Perimeter Institute for Theoretical Physics, 31 Caroline Street North, Waterloo, Ontario, N2L 2Y5, Canada}

\author{Hong-Ming Zhu}%email{hmzhu@bao.ac.cn}
\affiliation{Key Laboratory for Computational Astrophysics, National Astronomical Observatories, Chinese Academy of Sciences, 20A Datun Road, Beijing 100012, China}
\affiliation{University of Chinese Academy of Sciences, Beijing 100049, China}

\date{\today}
%\date{Received \today; published -- 00, 0000}

\begin{abstract}
We study the nonlinear $E$-mode clustering in Lagrangian space
by using large scale structure (LSS) $N$-body simulations
and use the displacement field information in Lagrangian space
to estimate the primordial linear density field.
We find that, compared to Eulerian nonlinear density fields,
the negative divergence of $E$-mode displacement fields
in Lagrangian space improves the cross-correlation with
initial density field by factor of 6 $\sim$ 7,
containing 2 orders of magnitude more primordial information. 
In reality, the displacement field is dominated by linear evolution
and insensitive of nonlinear shell-crossing, vorticity or
baryonic physics, and thus can be stably solved from only
the final matter distributions.
This illustrates ability of potential reconstruction algorithms,
to reconstruct baryonic acoustic oscillation (BAO) 
from current and future large scale structure surveys.

\end{abstract}

%\pacs{98.80.Es 95.36.+x}

\maketitle
\begin{figure}
 \centering
  \includegraphics[width=0.95\linewidth]{proj.pdf}
  \caption{Visualization of the nonlinear density field $\delta_N$ (top),
  linear density field $\delta_L$ (middle) and the raw reconstructed density
  field $\delta_R$ (bottom). These projections have 9.375 Mpc$/h$ thickness
  and 400 Mpc$/h$ per side. The top panel shows the nonlinear displacement
  field $\bs\Psi$ by the deformed mesh, which traces the LSS of $\delta_N$.}
  \label{fig.projection}
\end{figure}
\section{Introduction}\label{sec.intro}
Our universe starts from primordial Gaussian perturbations at a very early stage, 
and from those fluctuations, the gravitational instability drives the formation of 
the large scale structure (LSS) distribution of matter
\citep{1970A&A.....5...84Z,1985ApJ...292..371D}.
These structures grow 
linearly until the perturbations are large enough so that the first order 
perturbation theories are unable to analytically describe the LSS distributions 
\citep{2016JCAP...01..043M}. 
As a result, the final nonlinear LSS distribution contains higher order 
statistics, and thus makes it more challenging to be interpreted into basic
cosmological parameters. One such example is that, the baryonic acoustic oscillation (BAO)
scale can be used as a ``standard ruler'' to constrain the cosmic expansion history
and thus probes the dark energy properties \citep{2005NewAR..49..360E}, but
nonlinear evolution smears the BAO features and lowers the measurement
accuracy \citep{2005ApJ...633..560E,2012MNRAS.419.2949N}.
There are various attempts to recover earlier stages
of LSS, in which statistics are closer to Gaussian 
\citep{1992MNRAS.254..315W,2013MNRAS.436..759H}.
Because Gaussian fields can be adequately described by two-point statistics,
ideally after some recovery algorithms, more information can be extracted,
more straightforwardly, by power spectra or two-point correlation functions
\citep{2005MNRAS.360L..82R,2012MNRAS.421..832Y}.

Standard BAO reconstruction algorithms apply in Eulerian space.
They usually smooth the nonlinear density field on
linear scale ($\sim$10 Mpc$/h$) and reverse the large 
scale bulk flows by a negative Zel'dovich linear displacement
\citep{2007ApJ...664..675E,2009PhRvD..80l3501N,2009PhRvD..79f3523P}.
In the linear Lagrangian perturbation theory, the 
negative divergence of the the displacement field
$\bs\Psi(\bs q)$ respect to Lagrangian coordinates $\bs q$
gives the linear density field \citep{2010PhDT.........4J}.
Here we study the Lagrangian space clustering of dark matter
and use this information to estimate the initial density field.
Of course, the full displacement field $\bs\Psi(\bs q)$
is non-observable, as it requires the initial distributions of matter, however there are 
many techniques to estimate the displacement field from a final 
distribution of matter. For example, when a homogeneous initial matter distribution 
is assumed, there is a unique solution of curl-less displacement field to relate 
the initial and final distributions without shell-crossing. This solution can be 
solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}.
In 1-dimensional (1D) case, the exact solution 
simplifies to an ordering of matter elements by Eulerian coordinates.
Zhu {\it et al.} \cite{2016arXiv160907041Z} apply this algorithm to the result
of an 1D simulation \citep{2016JCAP...01..043M} and obtain an estimated
displacement field $\tilde\Psi(q)$, and find that this new method well recovers the linear
information and reconstructs 1D BAO peak in correlation function. 
In a bit more complicated 3D case, there are various
techniques to estimate $\tilde{\bs \Psi}(\bs q)$,
and we need to carefully consider effects of curl, shell-crossing in 3D cases.

Before these steps, we need to quantify the amount of linear information
that can be recovered from the full nonlinear displacement field $\bs\Psi(\bs q)$,
and further estimations $\tilde{\bs \Psi}(\bs q)$ can be compared with this result.
In this paper we run LSS $N$-body simulations and track the motion of
particles to obtain $\bs\Psi(\bs q)$. According to this we estimate the linear density
field and compare to the primordial density field of the initial
conditions. We describe the simulation and reconstruction algorithm in section
\ref{sec.method}, and we show the results in section \ref{sec.results}.
Discussion and conclusion are in section \ref{sec.discussion}.

\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{filter.pdf}
  \caption{Correlation functions $r(\delta_L,\delta_N)$ and $r(\delta_L,\delta_R)$
  (solid lines) and their scaled BAO damping models (dotted lines).}
  \label{fig.corr}
\end{figure}

\section{Method}\label{sec.method}
We show the LSS simulation and displacement field setups in section \ref{ss.sim}.
In section \ref{ss.reco}, we estimate the linear density field from the
displacement field $\bs\Psi(\bs q)$ from simulations.
Potential reconstruction algorithms are
based on an estimated displacement field $\tilde{\bs \Psi}(\bs q)$
instead of $\bs\Psi(\bs q)$. In the following sections we
simply call the estimation of linear density field
from $\tilde{\bs \Psi}(\bs q)$ ``reconstruction''.

\subsection{Simulation}\label{ss.sim}
We use the open source cosmological simulation code {\tt CUBE} %\footnote{github link of cafcube}
\citep{cafcube}.
Cosmological parameters are set as
$\Omega_m=0.27$, $\Omega_\Lambda=0.73$, $h_0=0.68$, $n_s=0.96$ and $\sigma_8=0.83$.
Initial conditions are generated at redshift $z=50$ 
using Zel'dovich approximation, and using a CLASS transfer function.
$N_p=512^3$ $N$-body particles are evolved via 
their mutual gravitational interactions to $z=0$, in a periodic box with $L=400$ 
Mpc$/h$ per side. The code is set to use standard a particle-mesh (PM) algorithm 
\cite{1988csup.book.....H} on a two-level mesh grids
(details see \cite{2013MNRAS.436..540H}) and cloud-in-cell
(CIC) is used in particle interpolations in force 
calculation and obtaining the density field $\rho({\bs x})$ in Eulerian coordinates 
${\bs x}$ at late stages. We use density contrast $\delta\equiv\rho/\lb\rho\rb-1$ 
to describe the density fluctuations. The primordial linear density field $
\delta_L$ is given by the initial stage and scaled to $z=0$ by the linear growth 
factor. In the top and middle panels of Fig.1 we show projections of the nonlinear density field
$\delta_N$ given by the simulation and the linear density field $\delta_L$.
$\delta_N$ is obtained by the particle distribution at redshift $z=0$, and
the particles are interpolated by using the cloud-in-cell (CIC) algorithm.
Because $\delta_N$ is highly nonlinear and follows an approximate
log-normal distribution, we plot $\log_{10}(\delta_N+1)$ instead.
The nonlinear evolution of $\delta_N$ makes it very different from $\delta_L$
in appearance.

The two-point statistics of these density fields are quantified by the cross power 
spectrum $P_{ij}(k)\equiv(2\pi)^{-3}\langle|\delta_i(k)||\delta_j(k)|\rangle$, 
where subscripts $i,j$ may refer to linear ($L$), nonlinear ($N$),
reconstructed ($R$), or noise ($n$) density 
fields. When $i=j$ it reduces to the auto power spectrum $P_{ii}(k)$ or $P(k)$. We 
usually plot the dimensionless power spectrum $\Delta^2(k)\equiv k^3P(k)/2\pi^2$.

\subsection{Reconstruction}\label{ss.reco}
In the simulation, we use particle-ID (PID) to record the initial (Lagrangian) location ${\bs 
q}$ of particles, and the information is tracked until the $z=0$ and we can get the 
Lagrangian displacement vector ${\bs \Psi}\equiv{\bs x}-{\bs q}$ for every 
particle. Then these vectors are interpolated onto the initial Lagrangian 
coordinates ${\bs q}$ of particles and we get the displacement field ${\bs \Psi}
({\bs q})$.
To visualize the $\bs\Psi$ field, we draw a 3D uniform mesh over the volume,
and use the given $\bs\Psi$ field to deform the mesh according to the direction
and physical amplitude of $\bs\Psi$. In the top panel of Fig.\ref{fig.projection},
The resulting mesh illustrates a pseudo
``curvilinear coordinate'' similar to \cite{1995ApJS..100..269P},
however the mesh can be overlapped due to shell-crossing. The densest mesh grids
trace the densest structures of $\delta_N$, whereas the undeformed grid positions
are the Lagrangian coordinates in which we do the reconstruction.
The raw reconstructed density field is given by the differential motion of matter 
elements,
\begin{equation}
    \delta_R=-\nabla\cdot{\bs \Psi}({\bs q}).
\end{equation}
Because the reconstruction processes are implemented on Lagrangian coordinates,
$\delta_R$ takes the coordinates of $\bs q$ instead of $\bs x$.
We just write $\bs q$'s Fourier wave number
$k_q$ as $k$ to simplify the expression.

\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{power.pdf}
  \caption{Power spectra of $\delta_L$, $\delta_N$, and the decomposition
  of $P_{LL}$ into correlated parts and noise terms according to equation
  (\ref{eq.power}).}
  \label{fig.recopower}
\end{figure}

\begin{figure}[t] \centering
  \includegraphics[width=1.0\linewidth]{pdfs.pdf}
  \caption{Probability distribution functions (PDFs)
  $f(\delta_L,\delta_N)$ and $f(\delta_L,\delta_R)$, showing in the upper
  and lower panels respectively. Both red curves on $x$-axes show $f(\delta_L)$, following
  a Gaussian distribution. The two curves on $y$-axes in upper and lower panel
  show respectively $f(\delta_N)$ and $f(\delta_R)$.}
  \label{fig.pdfs}
\end{figure}

To quantify the linear information in the reconstructed density field, we decompose $\delta_L$ in Fourier space as
\begin{equation}\label{eq.decompose}
    \delta_L(k)=r'\delta_R+\delta_n,
\end{equation}
where $r'\delta_R$ is completely correlated with linear density $\delta_L$. Correlating equation (\ref{eq.decompose}) with $\delta_R$ gives
\begin{equation}
    P_{LR}=r'P_{RR}+P_{nR},
\end{equation}
where $P_{ij}\equiv\langle\delta_i\delta_j\rangle$ denotes the cross power 
spectrum. Since $\delta_n$ is uncorrelated with $\delta_R$, $P_{nR}=0$. With the 
definition of cross correlation coefficient $r(\delta_L,\delta_R)\equiv P_{LR}/\sqrt{P_{LL}P_{RR}}$ 
and bias $b^2=P_{RR}/P_{LL}$, we solve $r'=P_{LR}/P_{RR}=rb^{-1}$.
Note that these computations above and below can also be applied on $\delta_N$ for
comparison, by replacing ``$_R$'' with ``$_N$'' in the equations,
while we do not rewrite them explicitly
in the paper for simplicity. From these, we plot the cross 
correlation coefficient $r_{LN}=r(\delta_L,\delta_N)$ and
$r_{LR}=r(\delta_L,\delta_R)$ in Fig.\ref{fig.corr}. Clearly, $
\delta_R$ contains much more linear information on smaller scales.

According to equation (\ref{eq.decompose}), the auto power spectrum is decomposed as
\begin{equation}\label{eq.power}
    P_{LL}=r^2b^{-2}P_{RR}+P_{nn},
\end{equation}
and $P_{nn}=(1-r^2)P_{LL}$. Then we construct a Wiener filter to filter out the uncorrelated part in $\delta_R$:
\begin{equation}
    W(k)=\frac{r^2b^{-2}P_{RR}}{r^2b^{-2}P_{RR}+P_{nn}}=r^2.
\end{equation}
The optimal reconstructed density is given by
\begin{equation}
    \tilde\delta_R=Wb^{-1}\delta_R,
\end{equation}
and the optimal reconstructed power spectrum is given by
\begin{equation}\label{eq.opt}
    \tilde P=W^2b^{-2}P_{RR}=W^2P_{LL}+W^2b^{-2}P_{NN}.
\end{equation}
Here $W^2$ describes the damping of the linear power spectrum.

\section{Results}\label{sec.results}

To visualize the above algorithms, a projection of $\delta_R$ is plotted in
the bottom panel of Fig.\ref{fig.projection}, which looks closer to 
$\delta_L$ compared to $\delta_N$. However the smallest scales are unable
to be reconstructed.

As discussed in section \ref{ss.reco},
Fig.\ref{fig.corr} shows the cross correlation functions
$r_{LN}$ and $r_{LR}$. The latter extends the correlation
with $\delta_L$ to smaller scales by nearly an order of
magnitude. The extra correlation scales well cover the BAO
scales of our interest.

In Fig.\ref{fig.recopower}, we show the auto power spectra of
$\delta_L$ and $\delta_N$ in black dashed and blue solid curves.
Their difference shows the nonlinear evolution of LSS on 
small scales. Their cross power (not shown for clarity of the figure)
drops to a very low value on scales $k\gtrsim 0.1h/$Mpc,
indicating a loss of linear information in the
nonlinear power spectrum $P_{NN}$. This scenario directly
leads to how $\P_{LL}$ is decomposed according to equation
(\ref{eq.power}). In nonlinear case (with ``$_R$'' replaced
by ``$_N$'' equation (\ref{eq.power})), on small
scales $k\gtrsim 0.1h/$Mpc, $P_{LL}$ is dominated by
uncorrelated, nonlinear noise, shown in the green
dotted line. In reconstructed case, however, $P_{LL}$
is decomposed into orange dash-dotted correlated
part and purple dotted uncorrelated part according to
equation (\ref{eq.power}). The correlated power spectrum
is dominated on BAO scales of our interest.

To quantify the improvement of cross-correlation in the power
spectrum, we compute the 
damping factors $W^2(k)$ respectively for the optimal filtered 
nonlinear and reconstructed density fields $\tilde\delta_N$
and $\tilde\delta_R$.
We fit Gaussian BAO damping models
${\mathcal D}(k)=\exp(-k^2\Sigma^2/2)$ to these $W^2(k)$'s
and give $\Sigma=1.8$ Mpc$/h$ and $\Sigma=12$ 
Mpc$/h$ for reconstructed and nonlinear fields.
Since ${\mathcal D}(k)=W^2=r^4$, we plot
${\mathcal D}_N^{1/4}$ and ${\mathcal D}_R^{1/4}$
over $r_{LN}$ and $r_{LR}$ in Fig.\ref{fig.corr}.
The analyses are repeated with various box sizes (100, 300, 800 Mpc$/h$ per side)
and give consistent results.

To further illustrate the improvement in real space
one point function correlations,
in Fig.\ref{fig.pdfs} we use the probability distribution
as functions (PDFs) of $(\delta_L,\delta_N)$ and $(\delta_L,\delta_R)$
to show the point-point correlation between these two pairs of density
fields. Since $\delta_n$ in equation (\ref{eq.decompose}) is
uncorrelated, we use Wiener filtered fields. To keep the
consistency over $\delta_L$, $\delta_N$ and $\delta_R$, we
use the $W(k)=r^2_{LR}$ as the Wiener filter.
The grey-scaled plots in the center of both panels show the 2-variable PDFs,
whereas their projections onto each variable are just
1-variable PDFs -- $f(\delta_L)$, $f(\delta_N)$ and $f(\delta_R)$, shown
as red/blue curves on the axes of Fig.\ref{fig.pdfs}.
In the top panel, $\delta_N$ shows an approximate log-normal
distribution (blue curve) and $\delta_L$ follows an expected
Gaussian distribution. They show tiny positive correlation in the
2D PDF.
Because in Fourier space, $\delta_L$ and $\delta_N$ have
correlations on only very large, linear scales (Fig.\ref{fig.corr}),
they result in little correlation in real space -- initial density fluctuations
in Lagrangian coordinates are evolved/transformed to Eulerian coordinates.
As the reconstruction is done in Lagrangian space, it recovers
certain amount of correlation, as shown in the 2D PDF of the bottom panel
of Fig.\ref{fig.pdfs}. One can also see that $\delta_R$ follows
a much closer Gaussian distribution (blue curve of the bottom panel).
We notice that the less-dense regions of $\delta_L$ are strongly
correlated with $\delta_R$, whereas denser regions of $\delta_L$ are
reconstructed less better, and $\delta_R$ shows more dispersion.
These second uncorrelated peaks in the 2D PDF and $f(\delta_R)$
are caused by nonlinear effects. We examine the reconstruction on
higher redshifts, and confirm that these second uncorrelated peaks
damp out as we go to higher redshifts.

\section{Discussion and conclusion}\label{sec.discussion}
We extract the actual displacement field of matter elements in cosmological $N$-body
simulations, and use this displacement field to study the LSS nonlinear
clustering in Lagrangian space. The displacement information is used to reconstruct
the primordial linear perturbations. The result shows a prominent improvement from
$r_{LN}$ to $r_{LR}$ in Fig.\ref{fig.corr} -- recovering the lost linear information on
nearly an order of magnitude smaller scales.
This is achieved by implementing differential movement information
of matter elements on Lagrangian coordinates, rather than on
Eulerian coordinates. This result illustrates the feasibility
of using estimated displacement field $\tilde{\bs \Psi}(\bs q)$ to reconstruct primordial linear
density field. A straightforward example of a estimation of $\tilde{\bs \Psi}(\bs q)$
is given by \cite{1995ApJS..100..269P,1998ApJS..115...19P}.
%\tcr{More complicated reconstruction algorithms, e.g., least action
%algorithms \cite{1989ApJ...344L..53P,2000ApJ...544...21G} can give
%the exact $\delta_L$}, where
%shell-crossing effects are also considered.
In reality, one needs
to consider all aspects including vorticity, shell-crossing, bias, noise
and data complexities. The impact of these factors can be quantitatively
compared with the impact of different estimation methods, and with
the exact solution by $N$-body simulations.

The advantage of using displacement field in reconstruction is
its insensitive response from highly nonlinearities --
densest regions form because matter elements, after shell crossing,
stop experiencing the linear extrapolated of Zel'dovich approximation.
i.e. the displacement fields are dominated by early stage linear processes,
which is the Lagrangian-Eulerian coordinate transform,
while late stage shell-crossing, nonlinear and baryonic processes
only fine-tunes the final position $\bs x$. In contrast, traditional
treatments in reconstruction deals directly on density fields which sensitively
relies on nonlinear processes --  density values can vary by orders
of magnitude due to nonlinear/baryonic physics and many sources of errors.

\section*{Acknowledgements}
We thank Pengjie Zhang for helpful discussions which triggered
the inspiration of the work.
HRY acknowledges General Financial Grant No.2015M570884 and Special Financial Grant No.2016T90009 from the China Postdoctoal Science Foundation.
HMZ acknowledges the support of the Chinese MoST 863 program under Grant
No. 2012AA121701, the CAS Science Strategic Priority Research Program
XDB09000000 and the NSFC under Grant No. 11373030.
The Dunlap Institute is funded through an endowment established by the David Dunlap family and the University of Toronto.
Research at the Perimeter Institute is supported by the Government of Canada
through Industry Canada and by the Province of Ontario through the Ministry of
Research $\&$ Innovation.

%\bibliographystyle{apsrev4-1}
\bibliographystyle{h-physrev3}
\bibliography{../haoran_ref}

\end{document}